---
title: "MVPAnalysis"
author: "Viren Gadkari"
date: "10/23/2021"
output: html_document
---

```{r}
library(tidyverse)


df <- read_csv('/Users/viren1/Documents/Data Science/SAAConference/MVPProject/mvp_final.csv')


df

```








```{r}

#df <- df %>% drop_na() %>% mutate(Year = str_remove(Year, ' NBA & ABA Awards Voting')) 

#df <- df %>% mutate(Year = str_remove(Year, ' NBA Awards Voting'))


df <- df %>% drop_na() %>% select(-c(X1, `Unnamed: 0` ))
#df %>% summarise(min_year = min(Year), max_year = max(Year))



df


```




```{r}

df$Year <- gsub("\\-.*","",df$Year)


era_1 <- df %>% filter(between(Year, '1972', '1979'))

era_2 <- df %>% filter(between(Year, '1980', '1994'))



era_3 <- df %>% filter(between(Year, '1995', '2013'))

era_4 <- df %>% filter(between(Year, '2013', '2019'))


era_1 %>% count(Year) %>% ggplot(aes(x=reorder(Year,n), y=n)) + geom_col(fill='purple') + coord_flip()

era_2 %>% count(Year) %>% ggplot(aes(x=reorder(Year,n), y=n)) + geom_col(fill='pink') + coord_flip()

era_3 %>% count(Year) %>% ggplot(aes(x=reorder(Year,n), y=n)) + geom_col(fill='magenta') + coord_flip()

era_4 %>% count(Year) %>% ggplot(aes(x=reorder(Year,n), y=n)) + geom_col(fill='orange') + coord_flip()
```




```{r}
library(tsibble)

era_1$Year <- as.Date(era_1$Year, '%Y')
era_2$Year <- as.Date(era_2$Year, '%Y')
era_3$Year <- as.Date(era_3$Year, '%Y')
era_4$Year <- as.Date(era_4$Year, '%Y')

era_1 <- era_1 %>% select(-c(Year, X1))
era_2 <- era_2 %>% select(-c(Year, X1))
era_3 <- era_3 %>% select(-c(Year, X1))
era_4 <- era_4 %>% select(-c(Year, X1))


df_subset <- df %>% select(-c( X1))

df_subset %>% summarise(min_year = min(Year), max_year = max(Year))
```

```{r}

library(GGally)
library(ggcorrplot)


df_subset_correlations <- df %>% select(-c(Player, Tm))

ggcorr(df_subset_correlations, method = c("everything", "pearson")) 



```

#convert to csv
```{r}
write.csv(df_subset_correlations, '/Users/viren1/Documents/Data Science/SAAConference/subset_correlations.csv')
```




```{r}
library(broom)
library(corrplot)

df_subset_correlations <- df_subset_correlations %>% select(-c(Year))

cor_plot <- cor(df_subset_correlations)
cor_plot
```




```{r}
library(tidymodels)

```


PCA



Era 1
```{r}
#ERA 1

era_1 <- era_1 %>% select(-c(First,Share))


era1_rec <- recipe(`Pts Won` ~ ., data=era_1) %>%
    update_role(Player, Tm, new_role='id') %>% 
    step_normalize(all_predictors()) %>%
    step_pca(all_predictors(), id='pca', num_comp=4)

era1_rec %>% prep() %>%
  tidy(id = "pca", type = "variance") %>% 
  dplyr::filter(terms == "percent variance") %>% 
  ggplot(aes(x = component, y = value)) + 
  geom_col(fill = "#b6dfe2") + 
  xlim(c(0, 5)) + 
  ylab("% of total variance")

era1_rec


```


```{r}

era1_prep <- prep(era1_rec)

tidied_era_1 <- tidy(era1_rec, 2)

era1_baked <- bake(era1_prep, era_1)




tidied_era_1

#tidied_era_1 %>%
  #filter(component %in% paste0("PC", 1:4)) %>%
  #mutate(component = fct_inorder(component)) %>%
  #ggplot(aes(value, terms, fill = terms)) +
  #geom_col(show.legend = FALSE) +
  #facet_wrap(~component, nrow = 1) +
  #labs(y = NULL)
```


PCA1 vs PCA2
```{r}

era1_baked %>%
  ggplot(aes(PC1, PC2, label = Player)) +
  geom_point(aes(color = Tm), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)

```



PCA1 vs. PCA3
```{r}

era1_baked %>%
  ggplot(aes(PC1, PC3, label = Player)) +
  geom_point(aes(color = Tm), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)


```



PCA1 vs. PCA4

```{r}

era1_baked %>%
  ggplot(aes(PC1, PC4, label = Player)) +
  geom_point(aes(color = Tm), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)

```


PC2 vs. PC3

```{r}

era1_baked %>%
  ggplot(aes(PC2, PC3, label = Player)) +
  geom_point(aes(color = Tm), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)

```


PC2 vs. PC4

```{r}
era1_baked %>%
  ggplot(aes(PC2, PC4, label = Player)) +
  geom_point(aes(color = Player), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)
```





PC3 vs. PC4
```{r}

era1_baked %>%
  ggplot(aes(PC3, PC4, label = Player)) +
  geom_point(aes(color = Player), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)

```



















```{r}
era_2 <- era_2 %>% select(-c(First, Share))

era2_rec <- recipe(`Pts Won` ~ ., data=era_2) %>%
  update_role(Player, Tm, new_role='id') %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), id='pca', num_comp = 4)

era2_prep <- prep(era2_rec)

tidied_era_2 <- tidy(era2_rec, 2)

tidied_era_2

#tidied_era_2 %>%
 # filter(component %in% paste0("PC", 1:4)) %>%
 # mutate(component = fct_inorder(component)) %>%
  #ggplot(aes(value, terms, fill = terms)) +
  #geom_col(show.legend = FALSE) +
  #facet_wrap(~component, nrow = 1) +
 # labs(y = NULL)

```



```{r}
era_3 <- era_3 %>% select(-c(First, Share))



```



```{r}
era_4 <- era_4 %>% select(-c(First, Share))



```



##DONT RUN THIS CHUNK

## RUN df
```{r}

df_subset_pca <- df %>% select(-c(`Pts Max`, Share, WS, First))
```




```{r}

library(tidymodels)




pca_rec <- recipe(`Pts Won`~., data=df_subset_pca)  %>%
  update_role(Player, Tm, Year, new_role='id') %>% 
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), id='pca', num_comp=4)

pca_rec %>% 
  prep() %>%
  tidy(id = "pca", type = "variance") %>% 
  dplyr::filter(terms == "percent variance") %>% 
  ggplot(aes(x = component, y = value)) + 
  geom_col(fill = "#b6dfe2") + 
  xlim(c(0, 5)) + 
  ylab("% of total variance")



pca_prep <- prep(pca_rec)


juice(pca_prep)
```



```{r}
tidied_pca <- tidy(pca_prep)

tidied_pca
```


```{r}
tidied_pca <- tidy(pca_prep, 2)

tidied_pca



tidied_pca %>%
  filter(component %in% paste0("PC", 1:4)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)
```

```{r}

library(tidytext)
tidied_pca %>%
  filter(component %in% paste0("PC", 1:4)) %>%
  group_by(component) %>%
  top_n(8, abs(value)) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(
    x = "Absolute value of contribution",
    y = NULL, fill = "Positive?"
  )
```





```{r}

pca_baked <- bake(pca_prep, df_subset_pca)

pca_baked

pca_baked$Year <- as.character(pca_baked$Year)





pca_baked %>% filter(between(Year, '1955', '1979')) %>%
  ggplot(aes(PC1, PC2, label = Player)) +
  geom_point(aes(color = `Pts Won`), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)  +
  scale_colour_gradient(low = "lightblue", high = "darkgreen", na.value = NA)
```





```{r}
pca_baked %>% filter(between(Year, '1980', '1994')) %>%
  ggplot(aes(PC1, PC2, label = Player)) +
  geom_point(aes(color = `Pts Won`), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL) +
  scale_colour_gradient(low = "lightblue", high = "darkgreen", na.value = NA)

```


```{r}
pca_baked %>% filter(between(Year, '1995', '2013')) %>%
  ggplot(aes(PC1, PC2, label = Player)) +
  geom_point(aes(color = `Pts Won`), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL) +
  scale_colour_gradient(low = "lightblue", high = "darkgreen", na.value = NA)


```

```{r}
pca_baked %>% filter(between(Year, '2013', '2019')) %>%
  ggplot(aes(PC1, PC2, label = Player)) +
  geom_point(aes(color = `Pts Won`), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL) +
  scale_colour_gradient(low = "lightblue", high = "darkgreen", na.value = NA)

```






```{r}
fit <- df_subset_pca %>% select(where(is.numeric)) %>% prcomp(scale=TRUE)



```

```{r}
fit %>%
  tidy(matrix = "rotation")


arrow_style <- arrow(
  angle = 20, ends = "first", type = "closed", length = grid::unit(8, "pt")
)

# plot rotation matrix
#fit %>%
  #tidy(matrix = "rotation") %>%
  #pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") %>%
  #ggplot(aes(PC1, PC2)) +
  #geom_segment(xend = 0, yend = 0, arrow = arrow_style) +
  #geom_text(
   # aes(label = column),
   # hjust = 1, nudge_x = -0.02, 
   # color = "#904C2F"
  #) +
  #xlim(-1.25, .5) + ylim(-.5, 1) +
  #coord_fixed() + # fix aspect ratio to 1:1
  #theme_minimal_grid(12)
```

```{r}
fit %>%
  tidy(matrix = "eigenvalues") %>%
  ggplot(aes(PC, percent)) +
  geom_col() +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.01))
  ) 
```



```{r}
library(ggbiplot)


ggbiplot(fit)
```




Age -- Player's age on February 1 of the season
Tm -- Team
Voting
First -- First Place Votes
Totals
G -- Games
Per Game
MP -- Minutes Played Per Game
PTS -- Points Per Game
TRB -- Total Rebounds Per Game
AST -- Assists Per Game
STL -- Steals Per Game
BLK -- Blocks Per Game
Shooting
FG% -- Field Goal Percentage
3P% -- 3-Point Field Goal Percentage
FT% -- Free Throw Percentage
Advanced
WS -- Win Shares
An estimate of the number of wins contributed by a player.
WS/48 -- Win Shares Per 48 Minutes
An estimate of the number of wins contributed by a player per 48 minutes (league average is approximately .100)




LASSO Regression

```{r}

#df_subset


df_subset_regression <- df %>% select(-c(`First`, `Share`, `Pts Max`, `WS`))

df_subset_regression
```


```{r}
#library(rstanarm)

#nba_glm <- stan_glm(`Pts Won` ~ ., data=df_subset, family=Gamma(link='log'))
```






```{r}
df_subset_regression

```




```{r}
nba_split <- initial_split(df_subset_regression)

nba_train <- training(nba_split)
nba_test <- testing(nba_split)




nba_recipe <- recipe(`Pts Won` ~ ., data=nba_train) %>%
  update_role(Player, Tm, Year, new_role='id') %>% 
  step_corr(all_numeric(), -all_outcomes()) %>%
  step_zv(all_numeric(), -all_outcomes()) %>%
  step_normalize(all_numeric(), -all_outcomes()) 


nba_prep <- nba_recipe %>% prep(strings_as_factors=FALSE)



lasso_model <- linear_reg(penalty=20, mixture=1) %>% set_engine('glmnet', intercept=FALSE, standardize=FALSE) 




ridge_model <- linear_reg(penalty=0.1, mixture=0) %>% set_engine('glmnet', intercept=FALSE, standardize=FALSE) 









wf <- workflow() %>% add_recipe(nba_recipe)




lasso_fit <- wf %>% add_model(lasso_model) %>% fit(data=nba_train)

lasso_fit %>% pull_workflow_fit() %>% tidy() %>% arrange(desc(estimate))# %>% ggplot(aes(x=reorder(term,estimate), y=estimate)) + geom_col(fill='blue') + coord_flip()


#lasso_gamma_fit <- wf %>% add_model(lasso_gamma_model) %>% fit(data=nba_train)

#lasso_gamma_fit %>% pull_workflow_fit() %>% tidy() %>% arrange(desc(estimate))

ridge_fit <- wf %>% add_model(ridge_model) %>% fit(data=nba_train)



#ridge_gamma_fit <- wf %>% add_model(ridge_gamma_model) %>% fit(data=nba_train)

#ridge_gamma_fit %>% pull_workflow_fit() %>% tidy() %>% arrange(desc(estimate))

ridge_fit %>% pull_workflow_fit() %>% tidy() %>% arrange(desc(estimate))# %>% ggplot(aes(x=reorder(term, estimate), y=estimate)) + geom_col(fill='red') + coord_flip()






#last_lasso_wf <- workflow() %>% add_model(lasso_model) %>% add_recipe(nba_recipe)


#last_lasso_fit <- last_lasso_wf %>% last_fit(nba_split)


#tidied_pca %>%
  #filter(component %in% paste0("PC", 1:4)) %>%
  #group_by(component) %>%
  #top_n(8, abs(value)) %>%
  #ungroup() %>%
 # mutate(terms = reorder_within(terms, abs(value), component)) %>%
 # ggplot(aes(abs(value), terms, fill = value > 0)) +
 # geom_col() +
 # facet_wrap(~component, scales = "free_y") +
 # scale_y_reordered() +
 # labs(
  #  x = "Absolute value of contribution",
  #  y = NULL, fill = "Positive?"
 # )




#nba_test$predictions <- lasso_test$.pred

#nba_test %>% rmse(truth=`Pts Won`, estimate=predictions)
```

```{r}
library(parsnip)
library(rpart)
library(rpart.plot)

nba_recipe_dt <- recipe(`Pts Won` ~ ., data=nba_train) %>%
  update_role(Player, Tm, Year, new_role='id') %>% 
  step_corr(all_numeric(), -all_outcomes()) 

decision_tree_model <- decision_tree(mode='regression') %>% set_engine('rpart')


decision_tree_wf <- workflow() %>% add_recipe(nba_recipe_dt)

decision_tree_fit <- decision_tree_wf %>% add_model(decision_tree_model) %>% fit(data=nba_train)

decision_tree_fit %>% pull_workflow_fit() 

rpart.plot(decision_tree_fit$fit$fit$fit)


```






```{r}
library(rpart.utils)


rules_df <- rpart.rules.table(decision_tree_fit$fit$fit$fit) %>%
  filter(Leaf==TRUE) %>%
  group_by(Rule) %>%
  summarise(Subrules = paste(Subrule, collapse=","))


df <- nba_train %>%
  mutate(Rule = row.names(decision_tree_fit$fit$fit$fit$frame)[decision_tree_fit$fit$fit$fit$where]) %>%
  left_join(rules_df, by="Rule")
head(df)


decision_tree_test <- decision_tree_fit %>% predict(nba_test) 

nba_test$predictions <- decision_tree_test$.pred

nba_test %>% rmse(truth=`Pts Won`, estimate=predictions)

```



```{r}
get_node <- function(rpart.object=decision_tree_fit$fit$fit$fit, data=nba_train, node.number=15) {
  data[which(decision_tree_fit$fit$fit$fit$where == node.number),]  
}
get_node()






```




```{r}

mvp_glm <- glm(`Pts Won` ~ Age + PTS + AST + `FG%` + `WS/48`, data=df_subset_regression, family=Gamma('identity'))

mvp_lm <- lm(`Pts Won` ~ Age + PTS + AST + `FG%` + `WS/48`, data=df_subset_regression)



```


```{r}
summary(mvp_glm)

plot(mvp_glm)

qqnorm(mvp_glm$residuals)
qqline(mvp_glm$residuals)



```


```{r}
summary(mvp_lm)

plot(mvp_lm)

hist(mvp_lm$residuals)
```






```{r}
ggplot(df_subset,aes(x=`Pts Won`)) + geom_histogram()
```


```{r}
nba_train_glmnet <- nba_train %>% select(-c(Player, Tm))
nba_test_glmnet <- nba_test %>% select(-c(Player, Tm))

x_train <- nba_train_glmnet 

y_train <- nba_train_glmnet %>% select(`Pts Won`)


x <- model.matrix(`Pts Won` ~ ., x_train)[,-1]
y <- x_train %>% select(`Pts Won`) %>% unlist() %>% as.numeric()
```


```{r}
df_subset_regression <- df_subset_regression %>% select(-c(Player, Tm))
```

```{r}

x <- model.matrix(`Pts Won` ~ ., df_subset_regression)[,-1]
y <- df_subset_regression %>% select(`Pts Won`) %>% unlist() %>% as.numeric()
grid = 10^seq(10, -2, length = 100)

x <- scale(x)

y <- scale(y)

train <- df_subset_regression %>% sample_frac(0.60)

test <- df_subset_regression %>% setdiff(train)


x_train <- model.matrix(`Pts Won` ~ ., train)[,-1]
x_test <- model.matrix(`Pts Won` ~ ., test)[,-1]

x_train <- scale(x_train)

x_test <- scale(x_test)

y_train <- train %>% select(`Pts Won`) %>% unlist() %>% as.numeric()

y_test <- test %>% select(`Pts Won`) %>% unlist() %>% as.numeric()

y_train <- scale(y_train)

y_test <- scale(y_test)

lasso_mod <- glmnet(x_train, y_train, alpha=0, lambda=10, standardize=FALSE)



```


```{r}


set.seed(1)
cv.out <- cv.glmnet(x_train, y_train,alpha=0, lambda=grid, standardize=FALSE)

plot(cv.out)


```





```{r}
bestlam <- cv.out$lambda.min
bestlam
#Qlasso_pred <- predict(lasso_mod, s=bestlam, newx=x_test)

#Qmean((lasso_pred-y_test)^2)
```

```{r}
out <- glmnet(x, y, alpha=0, lambda=grid, standardize=FALSE)

lasso_coef <- predict(out, type='coefficients', s=bestlam)

lasso_coef



summary(lasso_mod)
```

```{r}

lasso_coef[lasso_coef != 0]

```




Factor Analysis






Kmeans with PCA


```{r}
#df_subset_kmeans


pca_baked_kmeans <- pca_baked %>% select(-c(Player, Tm))



cls <- kmeans(pca_baked_kmeans, centers=3)
cls_2 <- kmeans(df_subset_regression, centers=3)


pca_baked$cluster <- as.character(cls$cluster)
df_subset_regression$cluster <- as.character(cls_2$cluster)

df_subset_regression$Player <- df$Player

df_subset_regression



```




```{r}
df_subset_regression %>%
  ggplot(aes(`WS/48`, `Pts Won`, label=Player)) +
  geom_point(aes(color = cluster), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL) 
```




```{r}
df_subset_regression %>%
  ggplot(aes(PTS, `Pts Won`, label=Player)) +
  geom_point(aes(color = cluster), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL) 
```


```{r}
ggplot(df_subset_regression, aes(x=AST, y=log(`Pts Won`))) + geom_point()
       
       
```







```{r}

library(psych)

```

```{r}

df_subset_correlations <- df_subset_correlations %>% select(-c(Share, First, `Pts Won`, `Pts Max`))


nba_pairwise_cor <- cor(df_subset_correlations, use= "pairwise.complete.obs")

```

```{r}
nba_cor <- cor(df_subset_correlations)

factors_data <- fa(r=nba_pairwise_cor, nfactors=4)



results <- data.frame(Factor_1 = factors_data[1])

#factors_data  %>% ggplot(aes(x=MR1, y=MR2, label(rownames(factors_data)))) + geom_point() +geom_text(check_overlap = TRUE, hjust = "inward") 

```



```{r}
fit <- factanal(df_subset_correlations, 3, 'cluster')

load <- fit$loadings[,1:2]
plot(load,type="n") # set up plot
text(load,labels=names(df_subset_correlations),cex=.7) # add variable names
```


```{r}
fit$loadings
```


```{r}

fa.diagram(factors_data)

```



```{r}


scree(nba_pairwise_cor, factors = FALSE)
```


```{r}
library(FactoMineR)
result <- PCA(df_subset_correlations)
```



A lot of gamma regression

```{r}

nba_train <- nba_train %>% select(-c(Player, Tm, WS))

#G is significant
#MP significant
#PTS significant
#TRB Significant
#AST SIGNificant
#STL sugnificant at 0.01
#BLK signifcant at 0.01
#FG% significant at 0.01
#3P% not significant
#FT% significant at 0.05
#WS/48 significant
#Age not significant
age_lm <- glm(`Pts Won` ~ Age , data=nba_train, family=Gamma('log'))


summary(age_lm)





```




Decision Tree

```{r}




```






